---
title: "Model 1: GBM"
author: "Junaira I. Ibrahim"
date: '2022-12-16'
output: pdf_document
---

*Importing packages*
```{r, warning=FALSE}
library(dplyr)    # for general data wrangling needs
library(gbm)      # for original implementation of regular and stochastic GBMs
library(h2o)      # for a java-based implementation of GBM variants
library(xgboost)  # for fitting extreme gradient boosting
library(caret)    # for classification and regression training
library(kernlab)  # for fitting SVMs
library(forcats)
library(bestNormalize)
library(vip)
library(rsample)
library(tidyverse)
library(caret)       # for general model fitting
library(rpart)       # for fitting decision trees
library(ROCR)
library(pROC)
```

*Importing the dataset*
The dataset used in this model is imported from `radiomics data`. It has 197 observations and 431 variables.
```{r}
datard <- read.csv("D:/MS_STATISTICS/STT225 Statistical Computing/FINAL PROJECT/radiomics_completedata.csv")
dim(datard)
```

*Checking for null and missing values*
```{r, results='hide'}
is.na(datard)
colSums(is.na(datard)) # no NA thus, there is no missing values
```

*Checking for normality*
```{r, warning=FALSE}
md1 = datard%>%select_if(is.numeric) 
datamd1 = lapply(md1[,-1], shapiro.test)
r = lapply(datamd1, function(x)x$p.value) #Extracting p-value only
s=unlist(r)    #to convert a list to vector
sum(s[s>0.05])
r$Entropy_cooc.W.ADC
```

Based on the results, there is only one variable who is normally distributed (i.e. Entropy_cooc.W.ADC). All the rest are not normally distributed. Hence, we will try to normalize the data using `orderNorm()` function.

*Normalizing the data*
```{r, warning=FALSE}
datard_norm = datard[,c(3,5:length(names(datard)))]
datard_norm = apply(datard_norm,2,orderNorm)
datard_norm = lapply(datard_norm, function(x) x$x.t)   #to transformed original data
datard_norm = datard_norm%>%as.data.frame()
```

Check the new data for normality
```{r, warning=FALSE}
datalr2 = lapply(datard_norm, shapiro.test)
r2 = lapply(datalr2, function(x) x$p.value)
s2 = unlist(r2)
sum(s2>0.05)
```
Based on the results, the rest of the variables is now normally distributed.

merging the normalized values into the original data, we have
```{r, warning=FALSE}
r3 = select(datard, c("Failure.binary",  "Entropy_cooc.W.ADC"))
datard_m = cbind(r3,datard_norm)
```

In this session, we will use `datard_n` and `datard_m`.

Getting the correlation of the whole data
```{r}
#correlation
newdatard = select(datard_m, -c("Failure.binary"))
cor.newdatard = cor(newdatard)
corr = round(cor.newdatard,2) # 2 decimals
heatmap(corr,Rowv=NA,Colv=NA,scale="none",revC = T)
```


```{r}
# for reproducibility
set.seed(123)  
```

The data `datard_m` is split into 80% of training data and 20% testing data.
```{r}
#80% training data - 20% testing data
rdsplit1 <- initial_split(datard_m, prop = 0.8, strata = "Failure.binary")
rdsplit1
rdtrain1 <- training(rdsplit1)
rdtest1  <- testing(rdsplit1)
```

#Gradient Boosting
GBMs build an ensemble of shallow trees in sequence with each tree learning and improving on the previous one. The name gradient boosting machine comes from the fact that this procedure can be generalized to loss functions other than SSE.

gbm has two training functions which are `gbm()` and `gbm.fit()`. The primary difference is that `gbm()` uses the formula interface to specify your model whereas `gbm.fit()`requires the separated $x$ and $y$ matrices. 
```{r}
gbmodel1 <- gbm(
  formula = Failure.binary ~ .,
  data = rdtrain1,
  distribution = "bernoulli",  # SSE loss function
  n.trees = 500,
  shrinkage = 0.1,
  n.minobsinnode = 10,
  cv.folds = 10
)
```


```{r}
# find index for number trees with minimum CV error
best <- which.min(gbmodel1$cv.error)
```


```{r}
# get MSE and compute RMSE
sqrt(gbmodel1$cv.error[best])
```
Model `gbmodel1` used the basic `gbm()` to train the model.This model has a minimum RMSE of 0.7612161.

Plotting the error curve
```{r}
# plot error curve
gbm.perf(gbmodel1, method = "cv")
```


```{r}
# create grid search
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  logloss = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {
  
  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = Failure.binary ~ .,
      data = rdtrain1,
      distribution = "bernoulli",
      n.trees = 500, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
    )
  })
  
  # add SSE, trees, and training time to results
  hyper_grid$logloss[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]
  
}
```


```{r}
# results
arrange(hyper_grid, logloss)
```
The results shows that a learning rate of 0.050 sufficiently minimizes our loss function (0.7716478) and requires 81 trees.


```{r}
# search grid
hyper_grid <- expand.grid(
  n.trees = 100,
  shrinkage = 0.01,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)

)
```


```{r}
# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Failure.binary ~ .,
    data = rdtrain1,
    distribution = "bernoulli",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))

}
```


```{r}
# perform search grid with functional programming
hyper_grid$logloss <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
  )
)

# results
arrange(hyper_grid, logloss)
```
The results shows that its logloss function 0.8786918 is greater than previous model.

```{r}
# refined hyperparameter grid
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "logloss",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)
```

```{r}
# perform grid search 
rdtrain1$Failure.binary=as.factor(rdtrain1$Failure.binary)
h2o.init()
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  y = "Failure.binary",
  training_frame = as.h2o(rdtrain1),
  hyper_params = hyper_grid,
  ntrees = 10,
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 5,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  stopping_metric="logloss",
  search_criteria = search_criteria,
  seed = 123
)
```


```{r}
# collect the results and sort by our model performance metric of choice
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "logloss", 
  decreasing = FALSE
)
```


```{r}
grid_perf
```


```{r}
# Grab the model_id for the top model, chosen by cross validation error
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let’s get performance metrics on the best model
h2o.performance(model = best_model, xval = TRUE)
```


```{r}
xgb_prep <- recipe(Failure.binary ~ ., data = rdtrain1) %>%
  step_integer(all_nominal()) %>%
  prep(training = rdtrain1, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Failure.binary")])
Y <- xgb_prep$Failure.binary
Y=as.numeric(Y)-1
```


```{r}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "binary:logistic",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)

```


```{r}
# minimum test CV RMSE
min(ames_xgb$evaluation_log$test_logloss_mean)
```


```{r}
# hyperparameter grid
hyper_grid <- expand.grid(
  eta = 0.01,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.5, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  logloss = 0,  
  trees = 0       
)

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 100,
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i], 
      lambda = hyper_grid$lambda[i], 
      alpha = hyper_grid$alpha[i]
    ) 
  )
  hyper_grid$logloss[i] <- min(m$evaluation_log$test_logloss_mean)
  hyper_grid$trees[i] <- m$best_iteration
}
```

```{r}
# results
hyper_grid %>%
  filter(logloss > 0) %>%
  arrange(logloss) %>%
  glimpse()
```

```{r}
# optimal parameter list
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)
```

```{r}
xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 3944,
  objective = "binary:logistic",
  verbose = 0
)
```

To print and plot the AUC values for training data, we use the following sytax in the chunks below.
```{r}
# Compute predicted probabilities on training data
m1_prob <- predict(xgb.fit.final, X, type = "prob")

# Compute AUC metrics for cv_model1,2 and 3 
perf1 <- prediction(m1_prob, rdtrain1$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")

# Plot ROC curves for cv_model1,2 and 3 
plot(perf1, col = "black", lty = 2)

# ROC plot for training data
roc(rdtrain1$Failure.binary ~ m1_prob, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)

xgb_prep <- recipe(Failure.binary ~ ., data = rdtest1) %>%
  step_integer(all_nominal()) %>%
  prep(training = rdtest1, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Failure.binary")])
```

plotting and AUC values for testing data
```{r}
# Compute predicted probabilities on testing data
m2_prob <- predict(xgb.fit.final, X, type = "prob")

# Compute AUC metrics for cv_model1,2 and 3 
perf2 <- prediction(m2_prob, rdtest1$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")

# Plot ROC curves for cv_model1,2 and 3 
plot(perf2, col = "black", lty = 2)

# ROC plot for training data
roc(rdtest1$Failure.binary ~ m2_prob, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)

```

we use `vip()` to construct a variable importance plot (VIP) of the top 20 features in the `bagging_2` model.
```{r}
# variable importance plot
vip(xgb.fit.final,num_features=20) 
```
